Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society

### 2023-12-07 manuscript/manuscript.qmd
Last-minute tweaks

---
title: "Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society"
short-title: "Enforcing Boundaries"
date: "2023-12-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  China's 2016 Overseas NGO (ONGO) Law is part of a larger global trend of increased legal restrictions on international nongovernmental organizations (INGOs). A growing body of research analyzes the broad effects of this crackdown on INGOs, finding a divergence in formal *de jure* laws and the *de facto* implementation of those laws. The causes and mechanisms of this divergence remain less explored. Why do authoritarian governments allow—and often collaborate—with some INGOs while harshly regulating or expelling others? What determines the openness of the practical legal operating environment for INGOs? In this paper, we use the case of China to explore how political demands to both restrict and embrace INGOs have shaped the international nonprofit sector in the five years since the ONGO Law came into effect. We argue that in an effort to bolster regime stability, governments use civil society laws as policy tools to influence INGO behavior. We find that INGO issue areas, missions, and pre-existing relationships with local government officials influence the degree of operating space available for INGOs. We test this argument with a mixed methods research design, combining Bayesian analysis of administrative data from all formally registered INGOs with a comparative case study of two environmental INGOs. Our findings offer insights into the practical effects of INGO restrictions and the dynamics of closing civic space worldwide.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: 7,984
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---

@@

```{r fig-issue-preds, fig.width=8, fig.height=6.5}
#| include: true
#| echo: false
#| column: body-outset
#| fig-cap: "Posterior predictions conditional on organization issue area. 
#|   **A**: Predicted proportion of INGOs working in 1 province, all 32 
#|   provinces, or somewhere in between. 
#|   **B**: Expectation of posterior predictive distributions."

>>>>

---
title: "Enforcing Boundaries: China's Overseas NGO Law and Operational Constraints for Global Civil Society"
short-title: "Enforcing Boundaries"
date: "2023-12-07"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
correspondence-prefix: "Correspondence concerning this article should be addressed to"
abstract: |
  China's 2016 Overseas NGO (ONGO) Law is part of a larger global trend of increased legal restrictions on international nongovernmental organizations (INGOs). A growing body of research analyzes the broad effects of this crackdown on INGOs, finding a divergence in formal *de jure* laws and the *de facto* implementation of those laws. The causes and mechanisms of this divergence remain less explored. Why do authoritarian governments allow—and often collaborate—with some INGOs while harshly regulating or expelling others? What determines the openness of the practical legal operating environment for INGOs? In this paper, we use the case of China to explore how political demands to both restrict and embrace INGOs have shaped the international nonprofit sector in the five years since the ONGO Law came into effect. We argue that in an effort to bolster regime stability, governments use civil society laws as policy tools to influence INGO behavior. We find that INGO issue areas, missions, and pre-existing relationships with local government officials influence the degree of operating space available for INGOs. We test this argument with a mixed methods research design, combining Bayesian analysis of administrative data from all formally registered INGOs with a comparative case study of two environmental INGOs. Our findings offer insights into the practical effects of INGO restrictions and the dynamics of closing civic space worldwide.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
word-count: "7,984"
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
toc-depth: 4
---

@@

```{r fig-issue-preds, fig.width=8.5, fig.height=7.5}
#| include: true
#| echo: false
#| column: body-outset
#| fig-cap: "Posterior predictions conditional on organization issue area. 
#|   **A**: Predicted proportion of INGOs working in 1 province, all 32 
#|   provinces, or somewhere in between. 
#|   **B**: Expectation of posterior predictive distributions."
